Expert-In-The-Loop Causal Discovery

### 2024-08-06 paper/main.tex
updates

% Your title must be in mixed case, not sentence case.
% That means all verbs (including short verbs like be, is, using,and go),
% nouns, adverbs, adjectives should be capitalized, including both words in hyphenated terms, while
% articles, conjunctions, and prepositions are lower case unless they
% directly follow a colon or long dash
\title{Expert-In-The-Loop Causal Discovery: Integrating Expert Knowledge}
\author{
    %Authors
    % All authors must be in the same font size and format.
    Written by AAAI Press Staff\textsuperscript{\rm 1}\thanks{With help from the AAAI Publications Committee.}\\
    AAAI Style Contributions by Pater Patel Schneider,
    Sunil Issar,\\
    J. Scott Penberthy,
    George Ferguson,
    Hans Guesgen,
    Francisco Cruz\equalcontrib,
    Marc Pujol-Gonzalez\equalcontrib
}
\affiliations{
    %Afiliations
    \textsuperscript{\rm 1}Association for the Advancement of Artificial Intelligence\\
    % If you have multiple authors and multiple affiliations
    % use superscripts in text and roman font to identify them.
    % For example,

@@

\begin{abstract}
	% TODO: Needs to be shortened.
	% In Pearl's framework, the initial step for performing a causal analysis
	% involves determining the causal structure between the variables from
	% data typically in the form of a Directed Acyclic Graph (DAG) or an
	% Structural Equation Model (SEM). This procedure, known as \emph{causal
	% discovery}, has been extensively studied for both DAGs and SEMs. While
	% numerous automated algorithms exist for estimating DAGs from datasets,
	% their application in applied fields remain limited, possibly due to a
	% lack of confidence in these algorithms stemming from them making
	% obvious mistakes, and the difficulties in choosing the right algorithm
	% for a given dataset. Consequently, in applied disciplines, these DAGs
	% are predominantly constructed manually based on domain expertise. Due
	% to being constructed by hand, it is vital that the fit of these models
	% are tested against data and the models are potentially modified to
	% acheive a better fit. On the other hand, in the field of SEMs models
	% are typically constructed manually through an iterative process of
	% assessing fit and modifying the model according to that. This process
	% is commonly known as \emph{Specification Search} where methods like
	% \emph{modification indices} combined with expert knowledge can be used
	% to guide this modification process. In the case of DAGs, although there
	% are tests to assess the global fit of a model by combining the tests
	% using implied conditional independencies of the model, but there are no
	% methods that can guide our modification process. In this paper, we
	% present a simple modification process aimed at helping researchers in
	% manually constructing or modifying DAGs outputted by an algorithm. This
	% modification process is based on using a (conditional) measure of
	% association between variables in the model to assess potential edges
	% that best explains unexplained correlation between variables which the
	% researcher specifies the direction of this edge. In the case of
	% continuous or discrete variables, the effect size measures of local
	% tests can be used as this measure of association, however for mixed
	% data there is no such measure. We present a measure of association
	% based on canonical correlations for mixed data. We also present a
	% graphical web-tool that can assist researchers in iteratively modifying
	% and constructing their DAGs. We theoretically show that in the presence
	% of an oracle, this iterative modification process is consistent in
	% recovering the true DAG. Emprically, we show that using the mixed data
	% measure of association, the modification process performs comparably to
	% PC and Hill-Climb Search algorithms if the user is able to correctly
	% identify the direction of the edge in one out of three cases.
\end{abstract}

>>>>

% Your title must be in mixed case, not sentence case.
% That means all verbs (including short verbs like be, is, using,and go),
% nouns, adverbs, adjectives should be capitalized, including both words in hyphenated terms, while
% articles, conjunctions, and prepositions are lower case unless they
% directly follow a colon or long dash
\title{Expert-In-The-Loop Causal Discovery}
\author{
    %Authors
    % All authors must be in the same font size and format.
    Written by AAAI Press Staff\textsuperscript{\rm 1}\thanks{With help from the AAAI Publications Committee.}\\
    AAAI Style Contributions by Pater Patel Schneider,
    Sunil Issar,\\
    J. Scott Penberthy,
    George Ferguson,
    Hans Guesgen,
    Francisco Cruz\equalcontrib,
    Marc Pujol-Gonzalez\equalcontrib
}
\affiliations{
    %Afiliations
    \textsuperscript{\rm 1}Association for the Advancement of Artificial Intelligence\\
    % If you have multiple authors and multiple affiliations
    % use superscripts in text and roman font to identify them.
    % For example,

@@

\begin{abstract}
	Causal Discovery has received substantial attention in the DAG
	literature and many automated algorithms have been proposed for
	constructing DAGs from data. However, the application of these
	algorithms in applied fields remain limited due to challenges in using
	them in practice for example they make obvious mistakes, output CPDAGs
	that need to be oriented using expert knowledge before estimation
	methods can be applied. As a result of these challenges, researchers
	typically construct these model by hand. In this paper, inspired by
	specification search in SEMs where the method suggests changes in the
	model that improves the fit the most, we present a method based on
	Conditional Independence testing and a novel measure of partial
	association. The method iteratively suggests addition and removal of
	edges between pairs of variables, while the user can choose the
	direction of the edge. Using empirical analysis we show that this
	manual construction approach can outperform automated algorithms if the
	user gets the direction correct in <x out of y> cases. We also show
	that the user expert can be replaced with an LLM.
\end{abstract}
